---
title: "Homework 3"
author: "Elise Mantell"
date: 2019-10-14
output: github_document
---
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(ggplot2)
library(ggridges)
```
### Problem 1

```{r upload dataset}
library(p8105.datasets) 
data("instacart") 

```

#### The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

```{r exploring the dataset}
str(instacart)
count(instacart, aisle) %>% ungroup() %>% arrange(desc(n))
```

##### _This dataset includes information about instacart orders, including items ordered, when they were ordred, and what aisle or department they come from. There are 1384617 observations of 15 variables. There are 134 aisles and the aisle with the most orders is the fresh vegetables aisle._

#### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r making number of orders plot} 
instacart %>%
count(aisle) %>% 
mutate(aisle = forcats::fct_reorder(aisle, n, .desc = TRUE)) %>% 
ggplot(aes(x = aisle, y = n)) + coord_flip() +
geom_bar(stat = "identity") +
labs(title = "Number of Orders in Aisle",
       x = "Aisle",
       y = "Number of Orders") 

```

##### _This is impossible to read because idk how to filter it yet._

#### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r making table of orders} 
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>% 
  top_n(3) %>% 
  arrange(desc(n)) %>%
  rename("Aisle" = aisle, "Product" = product_name,  "Number of Orders" = n) %>% 
  knitr::kable()
```

##### _The most commonly ordered packaged vegetables or fruits are all organic (spinach, raspberries, and blueberries). The second most commonly ordered dog food care item also appears to be organic. The most commonly ordered baking ingredients includes two types of sugar._


#### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r making table days} 
instacart %>% 
filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%  
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = round(mean(order_hour_of_day))) %>%
  spread(key = order_dow, value = mean_hour) %>% 
  rename("Product Name" = product_name) %>% 
  knitr::kable()
```

##### _People, on average, don't vary the time of day they order either apples or ice cream depending on the day of the week by more than about 3 hours._

### Problem 2

#### Focus on the “Overall Health” topic. Include only responses from “Excellent” to “Poor”. Organize responses as a factor taking levels ordered from “Poor” to “Excellent"

```{r loading and cleaning brfss}
brfss = p8105.datasets::brfss_smart2010 %>% 
janitor::clean_names() %>% 
rename(state = locationabbr, county = locationdesc) %>% 
filter(topic == "Overall Health") %>% 
mutate(response = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor")))
```

#### In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r states at 7+ locations in 2002}
brfss %>% 
  filter(year == 2002) %>% 
  distinct(state, county) %>% 
  count(state) %>% 
  filter(n == 7) 
```

##### _Three states were observed at more than 7 locations in 2002, two in the south (FL and NC) and one in the northeast (CT)._

```{r states at 7+ locations in 2010}
brfss %>% 
  filter(year == 2010) %>% 
  distinct(state, county) %>% 
  count(state) %>% 
  filter(n == 7) 
```

##### _Three different states were observed at more than 7 locations in 2010, one in the south (SC), one in the northeast (PA), and one in the west (CO)._


#### Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 



#### Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).


#### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State