---
title: "Homework 3"
author: "Elise Mantell"
date: 2019-10-14
output: github_document
---
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(ggplot2)
library(ggridges)
```
### Problem 1

```{r upload dataset}
library(p8105.datasets) 
data("instacart") 

```

#### The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

```{r exploring the dataset}
str(instacart)
count(instacart, aisle) %>% ungroup() %>% arrange(desc(n))
```

##### _This dataset includes information about instacart orders, including items ordered, when they were ordred, and what aisle or department they come from. There are 1384617 observations of 15 variables. There are 134 aisles and the aisle with the most orders is the fresh vegetables aisle._

#### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r fig.width=14, fig.height=12} 
instacart %>%
count(aisle) %>% 
filter(n > 10000) %>% 
mutate(aisle = forcats::fct_reorder(aisle, n, .desc = TRUE)) %>% 
ggplot(aes(x = aisle, y = n)) + coord_flip() +
geom_bar(stat = "identity") +
labs(title = "Number of Orders in Aisle",
       x = "Aisle",
       y = "Number of Orders")
```

##### _By far the aisles with the most orders are healthy (fruits, vegetables, and yogurt among the top 4)._

#### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r making table of orders} 
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>% 
  top_n(3) %>% 
  arrange(desc(n)) %>%
  rename("Aisle" = aisle, "Product" = product_name,  "Number of Orders" = n) %>% 
  knitr::kable()
```

##### _The most commonly ordered packaged vegetables or fruits are all organic (spinach, raspberries, and blueberries). The second most commonly ordered dog food care item also appears to be organic. The most commonly ordered baking ingredients includes two types of sugar._


#### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r making table days} 
instacart %>% 
filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  mutate(dow_name = recode(order_dow, 
                      '1' = "Mon",
                      '2' = "Tues",
                      '3' = "Wed",
                      '4'  = "Thurs",
                      '5' = "Fri",
                      '6' = "Sat",
                      '0' = "Sun")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = round(mean(order_hour_of_day))) %>%
  spread(key = order_dow, value = mean_hour) %>% 
  rename("Product" = product_name) %>% 
  knitr::kable()
```

##### _People, on average, don't vary the time of day they order either apples or ice cream depending on the day of the week by more than about 3 hours._

### Problem 2

#### Focus on the “Overall Health” topic. Include only responses from “Excellent” to “Poor”. Organize responses as a factor taking levels ordered from “Poor” to “Excellent"

```{r loading and cleaning brfss}
brfss = p8105.datasets::brfss_smart2010 %>% 
janitor::clean_names() %>% 
rename(state = locationabbr, county = locationdesc) %>% 
filter(topic == "Overall Health") %>% 
mutate(response = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor")))
```

#### In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r states at 7+ locations in 2002}
brfss %>% 
  filter(year == 2002) %>% 
  distinct(state, county) %>% 
  count(state) %>% 
  filter(n >= 7) 
```

##### _Six states were observed at more than 7 locations in 2002, two in the south (FL and NC) and the remaining in the northeast (CT, MA, NJ, PA)._

```{r states at 7+ locations in 2010}
brfss %>% 
  filter(year == 2010) %>% 
  distinct(state, county) %>% 
  count(state) %>% 
  filter(n >= 7) 
```

##### _Fourteen different states were observed at more than 7 locations in 2010, four in the south (FL, NC, SC, TX), the majority in the northeast (MA, MD, PA, NJ, NY, and PA), and the remaining in the west (CO, WA, OH, NE)._


#### Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 

```{r new dataset}
brfss_new = brfss %>% 
janitor::clean_names() %>% 
filter(response == "Excellent") %>% 
group_by(state, year) %>% 
summarise(avg_data_value = mean(data_value))
```

#### Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r SPAGHETTI}
brfss_new %>% 
  group_by(state, year) %>% 
  ggplot(aes(x = year, y = avg_data_value)) + 
    geom_line(aes(color = state)) + 
    labs(title = "Spaghetti Plot of Average Value Over Time for States",
         x = "Year", 
         y = "Average Value") + 
    theme(legend.position = "bottom", 
          legend.direction = "horizontal") + 
    guides(color = guide_legend(ncol = 17))
```

##### _This is terribly difficult to read with so many states. I have no meaningful comments, but it is quite pretty in my humble opinion._

#### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

```{r fig.width=8, fig.height=8}
brfss %>% 
  filter(state == "NY") %>%
  filter(year == "2006" | year == "2010") %>%
  select(year, data_value, county, response) %>%
  mutate(data_value = as.numeric(data_value), year = as.factor(year), county = as.factor(county)) %>%
  ggplot(aes(x = response, y = data_value)) + 
  geom_boxplot() + facet_grid(. ~year)
```

### Problem 3

#### Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).


```{r read and tidy accel}
accel_data = read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(day_type = ifelse(day == "Saturday" | day == "Sunday", "Weekend", "Weekday"))
```

##### _The dataset includes 35 observations of 1444 variables._

#### Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r aggregate activity}
accel_data_tot = 
accel_data %>%
mutate(tot_daily_activity = rowSums(.[4:1443]))
```

```{r making table}
accel_data_tot %>%  
group_by(week, day, tot_daily_activity) %>% 
select(week, day, tot_daily_activity) %>%
knitr::kable()
```

##### _There does not appear to be strong trends relating to either week number or day of the week._


#### Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.


```{r making hour variables}
accel_data_tot_hours = 
accel_data_tot %>%
mutate(H1 = rowSums(.[4:63])) %>%
mutate(H2 = rowSums(.[64:123])) %>%
mutate(H3 = rowSums(.[124:183])) %>%
mutate(H4 = rowSums(.[184:243])) %>%
mutate(H5 = rowSums(.[244:303])) %>%
mutate(H6 = rowSums(.[304:363])) %>%
mutate(H7 = rowSums(.[364:423])) %>%
mutate(H8 = rowSums(.[424:483])) %>%
mutate(H9 = rowSums(.[484:543])) %>%
mutate(H10 = rowSums(.[544:603])) %>%
mutate(H11 = rowSums(.[604:663])) %>%
mutate(H12 = rowSums(.[664:723])) %>%
mutate(H13 = rowSums(.[724:783])) %>%
mutate(H14 = rowSums(.[784:843])) %>%
mutate(H15 = rowSums(.[844:903])) %>%
mutate(H16 = rowSums(.[904:963])) %>%
mutate(H17 = rowSums(.[964:1023])) %>%
mutate(H18 = rowSums(.[1024:1083])) %>%
mutate(H19 = rowSums(.[1084:1143])) %>%
mutate(H20 = rowSums(.[1144:1203])) %>%
mutate(H21 = rowSums(.[1204:1263])) %>%
mutate(H22 = rowSums(.[1264:1323])) %>%
mutate(H23 = rowSums(.[1324:1383])) %>%
mutate(H24 = rowSums(.[1384:1443])) %>%
pivot_longer(H1:H24, 
  names_to = "hour", values_to = "aph") 
```




